---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```


# powergene: A simulation-based power analysis R Package to aid the design of robust semi-field vector control experiments

<!-- badges: start -->

<!-- badges: end -->

This R package simulates mosquito count data under short-term semi-field experimental designs, incorporating fixed treatment effects and random chamber variability. It’s built for pragmatic trial planning, MSc-level teaching, and reproducible research workflows.

## Installation

You can install the development version of powergene directly from
[GitHub](https://github.com/) with:


```r
# install.packages("devtools") if not already installed
install.packages("devtools")
devtools::install_github("akipingu/powergene")
```

```{r}
#load library
library(powergene)
```

## Example 1: Estimating power of a short-term SFE testing single intervention

This is a basic example which shows you how to apply the powergene package to estimate power of a short-term SFE

(i) Simulate a table of scenarios based on number of chamber per treatment provided, e.g., n.ch.per.trt=4. After specifying number of chamber per each treatment (control vs. intervention), the function sim.scen.shortsfe.sinint will output three columns including number of chambers per each treatment (replicates), treatment status (0 for control and 1 for intervention) and id for each individual chamber/replicate.

```{r}
sce.table <- sim.scen.shortsfe.sinint(n.ch.per.trt = 4)

#print the scenario table
sce.table
```


Simulate mosquitoes counts/abundance by specifying the number of chambers per treatment, e.g., n.ch.per.trt = 4, expected mosquitoes to be recaptured from a control chamber, e.g., lambda=50, intervention effect, e.g., interv.effect = 0.8 for 80% effect, and chamber-level variance, e.g., chamber.var = 0.1807. The resultant number of mosquitoes will be either with random effect included or without including the random effect.

```{r}
sim.mosquitoes <- sim.mosq.shortsfe.sinint(
  n.ch.per.trt = 4,
  lambda = 50,
  interv.effect = 0.8,
  chamber.var = 0.1807
)

#output mosquito counts per each chamber. I drop columns 4 and 6 because they are optional but you can print them if you want.
sim.mosquitoes <- sim.mosquitoes[, c(-4,-6)]
sim.mosquitoes
```

(iii) Estimate p-value using sim.pval.shortsfe.sinint function by specifying the number of chambers per treatment, e.g., n.ch.per.trt = 4, expected mosquitoes to be recaptured from a control chamber, e.g., lambda=50, intervention effect, e.g., interv.effect = 0.8 for 80% effect, and chamber-level variance, e.g., chamber.var = 0.1807.

```{r}
pvalue <- sim.pval.shortsfe.sinint(
    n.ch.per.trt = 4,
  lambda = 50,
  interv.effect = 0.8,
  chamber.var = 0.1807
)

#output the p-values
pvalue
```


(iv) Now you can estimate power by specifying the number of chambers per treatment, e.g., n.ch.per.trt = 4, expected mosquitoes to be recaptured from a control chamber, e.g., lambda=50, intervention effect, e.g., interv.effect = 0.8 for 80% effect, chamber-level variance, e.g., chamber.var = 0.1807, and the total simulations, e.g., nsim=100. A 100 simulations are just for practise, but it is recommended to run at least 1000 simulations if you are estimating power for a real experiment.

```{r}
power.estimate <- sim.power.shortsfe.sinint(n.ch.per.trt = 4,
  lambda = 50,
  interv.effect = 0.8,
  chamber.var = 0.1807, nsim = 100
  )

#print estimated power
power.estimate
```





